---
title: "R code for Data Science for Beginners"
subtitle: " Day 3: Individual Exercise"
author: Sara Beatty
date: "2024-9-11"  #r Sys.Date()
output:
  pdf_document: default
  html_document: default
---


## 1. Vectors 

### Create an object called `vec.a` which is a vector consisting of the numbers, 1, 3, 5, 7. You need to use the c function. 

vec.a <- c(1, 3, 5, 7)

### Create a vector called `vec.b` consisting of the numbers, 2, 4, 6, 8.

vec.b <- c(2, 4, 6, 8)

### Subtract vec.b from vec.a

vec.a - vec.b

### Create a new vector called vec.c by multiplying vec.a by vector vec.b

vec.c = (vec.a * vec.b)

### Create a new vector called vec.d by taking the square root of each member of vec.c

vec.d = sqrt(c(2, 12, 30, 56))

### What is the third element of the `vec.d` vector? Find out using square bracket. Note that since this is a vector, you only need to provide a single number inside the brackets. 

vec.d[3]


### Create a new vector called vec.e consisting of all the integers from 1 through 100. You should use the seq function, rather than writing down all the 100 integers individually. 

vec.e <- seq(1, 100)

### The mean function calculates the arithmetic mean of the numbers stored in an object. Using the mean function, calculate the mean of the `vec.e` vector.

mean(vec.e)

### As we saw in the joint exercise, the sum function calculates the sum of all the elements in an object. Calculate the sum of the `vec.e` vector. 

sum(vec.e)

### The length function returns the number of elements stored in an object. Using the length function, find the number of elements stored in the `vec.e` vector. 

length(vec.e)

### The mean of an object can be obtained by sum(X)/length(X) because the definition of the mean is the sum of elements divided by the number of elements. Now, using the sum and length functions, calculate the mean of the `vec.e` vector. Compare the answer with that obtained with the mean function

sum(vec.e) / length(vec.e)

We have learned that the by argument specifies an increment. For example, 

```{r}
seq(from = 0, to = 10, by = 2)
```

This creates a sequence that starts from 0 and ends with 10, and with an increment of 2. 


### Now, create a new object called olympic which is a sequence that starts from 1896 and ends with 2012, with an increment of 4. 

olympic <- seq(1896, 2012, by = 4)

### How many elements does the olympic vector contain? That is, what is the length of this vector? Find out by applying a function (not by manually counting the number of elements). 

length(olympic)


### So there are 30 elements in the olympic vector. Display all the  elements contained in the olympic vector. These are the years where olympic games were (supposed to be) held. Display the contents of the olympic vector. 

olympic

### Find out how many olympic games will have been held by the year 2400. Use the length and seq functions. 

length(seq(1896, 2400, by = 4))


## 2. Matrices 

### Create a new vector called `v1` consisting of the following numbers: 1, 3, 5, 7, 9, 11

v1 <- c(1, 3, 5, 7, 9, 11)

### Find out the length of this vector (Don't count the numbers by hand; use an appropriate function). 

length(v1)


### We will convert this vector into a matrix. That is, we will rearrange this vector so that it will have two dimensions (rows and columns). Since this vector has 6 numbers, if we want the matrix to have two rows, how many columns will there be?

There will be 3 columns.


### Create a matrix called mat.v using the following command: 

```{r}
# matrix(data = v1, nrow = 2)
```

matrix(data = v1, nrow = 2)

Take a look at the content of this matrix. How many columns are there? 


Notice how the numbers in vec.v are used to fill up the cells of `mat.v`.We can see that R did it "by column". That is, R first filled up the  first column of mat.v with the first two elements of vec.v, then moved on to the second and third columns. 


### You can use the `byrow` argument to change this. This argument takes one of two values, TRUE or FALSE (or T or F). That is, we write `matrix(data = v1, nrow = 2, byrow = TRUE)` Now, create an object called mat.w using the command above. 

mat.w <- matrix(data = v1, nrow = 2, byrow = TRUE)


Compare mat.v and mat.w. Do you see that R filled up the cells "by row" to create the mat.w matrix ? 

Many functions in R have arguments that take TRUE or FALSE like the byrow argument we just used. In most cases, functions have a default value. In the case of the matrix function, the default value for the byrow argument is FALSE, meaning that, if you don't specify anything, R will automatically sets byrow = FALSE. 



### Find the number in the second row, second column of `mat.w`

mat.w[2, 2]


### Find the number in the second row, second column of `mat.v`

mat.v[2, 2]




## 3. Lists 

### Create a list of months (as the names of the elements) with how many days each month has as the elements in the list

months <- list(
  January = 31,
  February = 28,
  March = 31,
  April = 30,
  May = 31,
  June = 30,
  July = 31,
  August = 31,
  September = 30,
  October = 31,
  November = 30,
  December = 31
)

### Display the number of days August has from the list 

months$August

### Convert the list to a vector

months.vector <- unlist(months)


## 4. Apply functions 

### Load R default data set mtcars 

data(mtcars)

### Use one of the apply functions to calculate the min value for each column/variable 
apply(mtcars, 2, min)

### Use one of the apply functions to indicate zero values in each column/variable 
apply(mtcars, 2, function(x) x == 0)


Finally, execute the entire contents of this file, making sure there is no error messages.

